---
title: "HW2"
author: "Andrew Li"
date: "2022-10-08"
output: github_document
---
https://github.com/P3BBL35/STAT433-HW2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(nycflights13)
library(ggplot2)
```

### What time of day should you fly if you want to avoid delays as much as possible?
```{r}
flights %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(hour) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  arrange(desc(proportion_arrived_on_time))
```
To avoid delays as much as possible, you want to fly at 7am, as it has the
highest proportion of flights that arrive on time.

### Does this choice depend on anything? Season? Weather? Airport? Airline? Find three patterns (“null results” are ok!)
The choice does seem to depend on weather and the airline. It does not seem to 
depend on the airport from which a flight took off.

#### Airline
```{r}
flights2 <- flights %>%
  select(year:day, hour, tailnum, carrier, arr_delay)
flights2 %>%
  left_join(airlines, by = 'carrier') %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(name) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  arrange(desc(proportion_arrived_on_time))
```

With the above chart, you can see that the proportions vary quite widely across
the different airlines at 7am. For example, AirTran Airways has the lowest proportion
of its flight making it on time at 0.532, where Alaska Airlines has a relatively
larger proportion, sitting at 0,819. However, we still need to take into account
the number of flights being recorded, as a lower sample size means a less accurate
record of their true average ability of being on time. For example, Frontier Airlines
has a perfect record, but only has 1 recorded flight at 7am.

#### Weather
```{r}
flights2 <- flights %>%
  select(year:day, hour, arr_delay)
flights2 %>%
  left_join(weather, by = c('year', 'month', 'day', 'hour')) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(hour) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n(),
            average_precip = mean(precip, na.rm = T)) %>%
  ggplot(aes(x = average_precip, y = proportion_arrived_on_time)) + geom_point()
```

There does not seem to be a very clear relationship between amount of precipitation
and the proportion of flights that arrived on-time, especially after removing the
outlier at (0,0)

```{r}
flights2 <- flights %>%
  select(year:day, hour, arr_delay)
flights2 %>%
  left_join(weather, by = c('month', 'day', 'hour')) %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(visib) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  ggplot(aes(x = visib, y = proportion_arrived_on_time)) +
    geom_point()
```

However, there is a clear relationship between the visibility due to weather conditions
and the proportion of flights that arrived on-time. And this makes sense. Lower
visibility makes it harder for planes to navigate in the air, causing them to
go through turbulence or go slower due to requiring navigational signals to make
sure they're going the right way/not going to crash into anything else. Thus, a
good way to increase your chances of being on-time is to ensure weather will be
clear at the time of takeoff.

#### Airport
```{r}
flights2 <- flights %>%
  select(year:day, hour, origin, arr_delay)
flights2 %>%
  left_join(airports, by = c('origin' = 'faa')) %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(name) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n())
```

There is no clear relationship between the airport from which you started flying
and the proportion of flights that arrive on time. Across all the airports, the
proportion of flights that arrived on time when departing from the airport sits
around 0.55-0.60. This means that if you're planning on taking a flight from New
York to anywhere else, the airport you choose does not really matter with regards
to being on time.
